### 2016-06-21 Kokkos_PG_Compilation.tex
Some updates

@@

\hline    
    KOKKOS\_ARCH (IN): & What backend Architecture to build for. \\
      \hspace{0.5cm}Options: & KNC,SNB,HSW,Kepler,Kepler30,Kepler35,Kepler37,Maxwell,
               Maxwell50,ARMv8,BGQ,Power7,Power8 \\
      \hspace{0.5cm}Default: & None -- this means no particular architecture flags
               are set. \\

@@

\hline
   KOKKOS\_CUDA\_OPTIONS (IN): & Enable optional settings specific to CUDA. \\
      \hspace{0.5cm}Options: & force\_uvm,use\_ldg,rdc \\
      \hspace{0.5cm}Default: & \\

>>>>

In order to compile Kokkos a C++11 compliant compiler is needed. 
For an up to date list of compilers which are tested on a nightly basis, please refer to the README on the github repository. 
At the time of writing supported compilers include: 
\begin{lstlisting}
  GCC   4.7.2, 4.8.x, 4.9.x, 5.1.x, 5.3.x
  Intel 14.x, 15.x, 16.x, 17.x 
  Clang 3.5.1, 3.6.x, 3.7.x, 3.8.x, 3.9.x; 
  Cuda  6.5, 7.0, 7.5, 8.0; 
  XL 13.3;
\end{lstlisting}

@@

\hline    
    KOKKOS\_ARCH (IN): & What backend Architecture to build for. \\
      \hspace{0.5cm}Options: & KNL,KNC,SNB,HSW,Kepler,Kepler30,Kepler35,Kepler37,Maxwell,
               Maxwell50,ARMv8,BGQ,Power7,Power8 \\
      \hspace{0.5cm}Default: & None -- this means no particular architecture flags
               are set. \\

@@

\hline
   KOKKOS\_OPTIONS (IN): & Enable optional settings \\
      \hspace{0.5cm}Options: & aggressive\_vectorization \\
      \hspace{0.5cm}Default: & \\

\hline
   KOKKOS\_CUDA\_OPTIONS (IN): & Enable optional settings specific to CUDA. \\
      \hspace{0.5cm}Options: & force\_uvm,use\_ldg,rdc,enable\_lambda \\
      \hspace{0.5cm}Default: & \\

### 2016-06-21 Kokkos_PG_HierarchicalParallelism.tex
Some updates

Instead of mapping a 1-D range of indices to hardware resources,
Kokkos' thread teams map a 2-D index range.
The first index is the \emph{league}, the index of the team.
The second index is the thread index within a team.
In CUDA this is equivalent to launching a 1-D grid of 1-D blocks.
The league size is arbitrary -- that is, it is only limited by the integer size type -- while the team size must fit in the hardware constraints.
As in CUDA, only a limited number of teams are actually active at the same time,
and they must run to completion before new ones are executed. 
Consequently it is not valid to use inter thread-team synchronization mechanisms
such as waits for events initiated by other thread teams.

@@

Kokkos exposes use of thread teams with the \lstinline!Kokkos::TeamPolicy! execution policy.
To use thread teams you need to create a \lstinline|Kokkos::TeamPolicy| instance.
It can be created inline for the parallel dispatch call.
The constructors requires two arguments: a league size and a team size. 
As with the  \lstinline|Kokkos::RangePolicy| a specific execution tag and a specific execution space can be given as optional template arguments.
\begin{lstlisting}
// Using default execution space and launching 
// a league with league_size teams with team_size threads each
Kokkos::TeamPolicy<> 
        policy( league_size, team_size );

// Using  a specific execution space to 
// run a n_worksets x team_size parallelism
Kokkos::TeamPolicy<ExecutionSpace> 
        policy( league_size, team_size );

@@

typedef TeamPolicy<ExecutionSpace>::member_type member_type;
// Create an instance of the policy
TeamPolicy<ExecutionSpace> policy (league_size, team_size);
// Launch a kernel
parallel_for (policy, KOKKOS_LAMBDA (member_type team_member) {
    // Calculate a global thread id
    int k = team_member.league_rank () * team_member.team_size () +
            team_member.team_rank ();
    // Calculate the sum of the global thread ids of this team
    int team_sum = team_member.reduce (k);
    // Atomicly add the value to a global value
    a() += team_sum;
  });
\end{lstlisting}

@@

Kokkos exposes scratch pads through a special memory space associated with the execution space:
\lstinline|execution_space::scratch_memory_space|.
You may allocate a chunk of scratch memory through the \lstinline|TeamPolicy| member type.
You may request multiple allocations from scratch, up to a user-provided maximum. 
The maximum is provided either through a function in the functor which returns a potentially team-size dependent value, 
or it can be specified as an argument to the constructor of the TeamPolicy. 
It is not valid to provide both values at the same time. 
The argument to the TeamPolicy can be used to set the shared memory size when using functors. 
One restriction on shared memory allocations is that they can not be freed during the lifetime of the team. 
This avoids the complexity of a memory pool,
and reduces the time it takes to obtain an allocation
(which currently is a few tens of integer operations to calculate the offset).

\begin{lstlisting}
template<class ExecutionSpace>
struct functor {
  typedef ExecutionSpace execution_space;
  typedef execution_space::member_type member_type;

@@

@@

You may use any of the three parallel patterns -- for, reduce, or scan -- at each level.
You may nest them and use them in conjunction with code that is aware of the league and team rank.
The different layers are accessible via special execution policies:
\lstinline|TeamThreadLoop| and \lstinline|ThreadVectorLoop|.

\subsection{Team and vector loops}\label{SS:Hierarchical:Nested:Loops}

The first nested level of parallel loops splits an index range over the threads of a team. 
This motivates the policy name \lstinline|TeamThreadRange|, 
which indicates that the loop is executed once by the team with the index range split over threads.
The loop count is not limited to the number of threads in a team, and how the index range is mapped to threads is architecture dependent.
It is not legal to nest multiple parallel loops using the \lstinline!TeamThreadRange! policy.
However, it is valid to have multiple parallel loops using the \lstinline!TeamThreadRange! policy follow each other in sequence, in the same kernel.  
Note that it is not legal to make a write access to POD data outside of the closure of a nested parallel layer. 
This is a conscious choice to prevent difficult to debug issues related to thread private, team shared and globally shared variables. 
A simple way to enforce this is by using the ``capture by value'' clause with lambdas. 
With the lambda being considered as \lstinline|const| inside the \lstinline!TeamThreadRange! loop,
the compiler will catch illegal accesses at compile time as a \lstinline|const| violation.

@@

@@

Kokkos provides the \lstinline|Kokkkos::single(Policy,Lambda)| function for this case.
It currently accepts two policies:
\begin{itemize}
\item \lstinline|Kokkos::PerTeam| restricts execution of the lambda's
  body to once per team
\item \lstinline|Kokkos::PerThread| restricts execution of the
  lambda's body to once per thread (that is, to only one vector lane
  in a thread)
\end{itemize}
The \lstinline!single! function takes a lambda as its second argument.
That lambda takes zero arguments, 
so its body must perform side effects in order to have an effect.
It must always be correct for the lambda to capture variables by value
(\lstinline![=]!, not \lstinline![&]!).
Thus, if the lambda captures by reference,
it must \emph{not} modify variables that it has captured by reference.

>>>>

Instead of mapping a 1-D range of indices to hardware resources,
Kokkos' thread teams map a 2-D index range.
The first index is the \emph{league rank}, the index of the team.
The second index is the \emph{team rank}, the thread index within a team.
In CUDA this is equivalent to launching a 1-D grid of 1-D blocks.
The league size is arbitrary -- that is, it is only limited by the integer size type -- while the team size must fit in the hardware constraints.
As in CUDA, only a limited number of teams are actually active at the same time,
and they must run to completion before new ones are executed. 
Consequently it is not valid to use inter thread-team synchronization mechanisms
such as waits for events initiated by other thread teams.

@@

Kokkos exposes use of thread teams with the \lstinline!Kokkos::TeamPolicy! execution policy.
To use thread teams you need to create a \lstinline|Kokkos::TeamPolicy| instance.
It can be created inline for the parallel dispatch call.
The constructors requires two arguments: a league size and a team size. 
In place of the team size a user can utilize \lstinline|Kokkos::AUTO| to let Kokkos guess a good team size for a given architecture.
Doing that is the recommend way for most developers to utilize the \lstinline|TeamPolicy|.
As with the  \lstinline|Kokkos::RangePolicy| a specific execution tag and a specific execution space can be given as optional template arguments.
\begin{lstlisting}
// Using default execution space and launching 
// a league with league_size teams with team_size threads each
Kokkos::TeamPolicy<> 
        policy( league_size, team_size );

// Using  a specific execution space to 
// run a n_worksets parallelism with Kokkos choosing the team size
Kokkos::TeamPolicy<ExecutionSpace> 
        policy( league_size, Kokkos::AUTO() );

@@

typedef TeamPolicy<ExecutionSpace>::member_type member_type;
// Create an instance of the policy
TeamPolicy<ExecutionSpace> policy (league_size, Kokkos::AUTO() );
// Launch a kernel
parallel_for (policy, KOKKOS_LAMBDA (member_type team_member) {
    // Calculate a global thread id
    int k = team_member.league_rank () * team_member.team_size () +
            team_member.team_rank ();
    // Calculate the sum of the global thread ids of this team
    int team_sum = team_member.reduce (k);
    // Atomicly add the value to a global value
    a() += team_sum;
  });
\end{lstlisting}

@@

Kokkos exposes scratch pads through a special memory space associated with the execution space:
\lstinline|execution_space::scratch_memory_space|.
You may allocate a chunk of scratch memory through the \lstinline|TeamPolicy| member type.
You may request multiple allocations from scratch, up to a user-provided maximum aggregate size. 
The maximum is provided either through a \lstinline|team_shmem_size| function in the functor which returns a potentially team-size dependent value, 
or it can be specified through a setting of the TeamPolicy \lstinline|set_scratch_size|. 
It is not valid to provide both values at the same time. 
The argument to the TeamPolicy can be used to set the shared memory size when using functors. 
One restriction on shared memory allocations is that they can not be freed during the lifetime of the team. 
This avoids the complexity of a memory pool,
and reduces the time it takes to obtain an allocation
(which currently is a few tens of integer operations to calculate the offset).

The following is an example of using the functor interface:
\begin{lstlisting}
template<class ExecutionSpace>
struct functor {
  typedef ExecutionSpace execution_space;
  typedef execution_space::member_type member_type;

@@

The \lstinline|set_scratch_size| function of the \lstinline|TeamPolicy| takes two or three arguments.
The first argument specifies the level in the scratch hierarchy for which a specific size is requested. 
Different levels have different restrictions. 
Generally the first level is restricted to a few tenths of kilobyte roughly corresponding to L1 cache size. 
The second level can be used to get an aggregate over all teams of a few gigabyte, corresponding to available 
space in high-bandwidth memory. 
The third level generally falls back to capacity memory in the node. 
The second and third argument are either per-thread or per-team sizes for scratch memory.
Note like previously discussed, the setter function does not modify the instance it is called on, but returns 
a copy of the policy object with adjusted scratch size request.

Here are some examples:
\begin{lstlisting}
TeamPolicy<> policy_1 = TeamPolicy<>(league_size, team_size).
                          set_scratch_size(1, PerTeam(1024), PerThread(32));
TeamPolicy<> policy_2 = TeamPolicy<>(league_size, team_size).
                          set_scratch_size(1, PerThread(32));
TeamPolicy<> policy_3 = TeamPolicy<>(league_size, team_size).
                          set_scratch_size(1, PerTeam(1024));
\end{lstlisting}

The total amount of scratch space available for each team will be the per-team value plus the per-thread value multiplied by the team-size. 
The interface allows users to specify those settings inline:
\begin{lstlisting}
parallel_for(TeamPolicy<>(league_size, team_size).set_scratch_size(1, PerTeam(1024)),
  KOKKOS_LAMBDA (const TeamPolicy<>::member_type& team) {
    ...
});
\end{lstlisting}

@@

You may use any of the three parallel patterns -- for, reduce, or scan -- at each level\footnote{The parallel scan operation is not implemented for all execution spaces on the thread level, and it doesn't support a TeamPolicy on the top level.}.
You may nest them and use them in conjunction with code that is aware of the league and team rank.
The different layers are accessible via special execution policies:
\lstinline|TeamThreadLoop| and \lstinline|ThreadVectorLoop|.

\subsection{Team loops}\label{SS:Hierarchical:Nested:Loops:Team}

The first nested level of parallel loops splits an index range over the threads of a team. 
This motivates the policy name \lstinline|TeamThreadRange|, 
which indicates that the loop is executed once by the team with the index range split over threads.
The loop count is not limited to the number of threads in a team, and how the index range is mapped to threads is architecture dependent.
It is not legal to nest multiple parallel loops using the \lstinline!TeamThreadRange! policy.
However, it is valid to have multiple parallel loops using the \lstinline!TeamThreadRange! policy follow each other in sequence, in the same kernel.  
Note that it is not legal to make a write access to POD data outside of the closure of a nested parallel layer. 
This is a conscious choice to prevent difficult to debug issues related to thread private, team shared and globally shared variables. 
A simple way to enforce this is by using the ``capture by value'' clause with lambdas, 
but ``capture by reference'' is recommended for release builds since it typically results in better performance.
With the lambda being considered as \lstinline|const| inside the \lstinline!TeamThreadRange! loop,
the compiler will catch illegal accesses at compile time as a \lstinline|const| violation.

@@

\subsection{Vector loops}\label{SS:Hierarchical:Nested:Loops:Vector}

At the innermost level of nesting parallel loops in a kernel is comprised of the \emph{vector}-loop.
Vector level parallelism works identical to the team level loops using the execution policy \lstinline|ThreadVectorRange|.
In contrast to the team-level, there is no legal way to exploit the vector level outside of a parallel pattern using the \lstinline|ThreadVectorRange|. 
However one can use such a parallel construct in- and outside of a \lstinline|TeamThreadRange| parallel operation.

\begin{lstlisting}
using Kokkos::parallel_reduce;
using Kokkos::TeamPolicy;
using Kokkos::TeamThreadRange;
using Kokkos::ThreadVectorRange;
parallel_for (TeamPolicy<> (league_size, team_size),
                 KOKKOS_LAMBDA (member_type team_member) {

    int k = team_member.team_rank();
    // The default reduction uses Scalar's += operator
    // to combine thread contributions.
    Scalar sum;
    parallel_reduce (ThreadVectorRange (team_member, loop_count),
      [=] (int& i, Scalar& lsum) {
        // ...
        lsum += ...;
      }, sum);

    parallel_for (TeamThreadRange (team_member, workset_size),
      [&] (int& j) {
      // You may provide a custom reduction as another
      // lambda together with an initialization value.
      Scalar product;
      Scalar init_value = 1;
      parallel_reduce (ThreadVectorRange (team_member, loop_count),
        [=] (int& i, Scalar& lsum) {
          // ...
          lsum *= ...;
        }, product, [=] (Scalar& lsum, Scalar& update) {
          lsum *= update;
        }, init_value);
      });
  });
\end{lstlisting}

As the name indicates the vector-level must be vectorizable. 
The parallel patterns will exploit available mechanisms to encourage vectorization by the compiler. 
When using the Intel compiler for example, the vector level loop will be internally decorated with 
\lstinline|#pragma ivdep|, telling the compiler to ignore assumed vector dependencies.

@@

Kokkos provides the \lstinline|Kokkkos::single(Policy,Lambda)| function for this case.
It currently accepts two policies:
\begin{itemize}
\item \lstinline|Kokkos::PerTeam| restricts execution of the lambda's
  body to once per team
\item \lstinline|Kokkos::PerThread| restricts execution of the
  lambda's body to once per thread (that is, to only one vector lane
  in a thread)
\end{itemize}
The \lstinline|single| function takes a lambda as its second argument.
That lambda takes zero arguments, 
so its body must perform side effects in order to have an effect.
It must always be correct for the lambda to capture variables by value
(\lstinline|[=]|, not \lstinline|[&]|).
Thus, if the lambda captures by reference,
it must \emph{not} modify variables that it has captured by reference.

### 2016-06-21 Kokkos_PG_ParallelDispatch.tex
Some updates

Kokkos lets users choose whether to use a functor or a lambda.
Lambdas are convenient for short loop bodies.
For a much more complicated loop body,
you might find it easier for testing to separate it out and name it as a functor.
Lambdas by definition are ``anonymous functions,'' meaning that they have no name.
This makes it harder to test them.
Furthermore, if you would like to use lambdas with CUDA,
you must have a sufficiently new version of CUDA (at the time of writing Lambdas are not supported yet supported by the release 7.0).
Finally, the ``execution tag'' feature, which lets you put together several different parallel loop bodies into a single functor,
only works with functors.  (See Chapter \ref{C:Hierarchical} for details.)

>>>>

Kokkos lets users choose whether to use a functor or a lambda.
Lambdas are convenient for short loop bodies.
For a much more complicated loop body,
you might find it easier for testing to separate it out and name it as a functor.
Lambdas by definition are ``anonymous functions,'' meaning that they have no name.
This makes it harder to test them.
Furthermore, if you would like to use lambdas with CUDA,
you must have a sufficiently new version of CUDA. 
At the time of writing CUDA 7.5 supports lambdas using a special flag. 
To enable this support, use the \lstinline|KOKKOS_CUDA_OPTIONS=enable_lambda| option.

Finally, the ``execution tag'' feature, which lets you put together several different parallel loop bodies into a single functor,
only works with functors.  (See Chapter \ref{C:Hierarchical} for details.)

### 2016-06-21 Kokkos_PG_Views.tex
Some updates

@@

Other layouts are possible.  For example, Kokkos has a ``tiled''
layout, where a tile's entries are stored contiguously (in either row-
or column-major order) and tiles have compile-time dimensions.  One
may also use Kokkos to implement Morton ordering or variants thereof.

@@

// Get 'LDA' for BLAS / LAPACK
int strides[2]; // any integer type works in stride()
A.stride (strides);
const int LDA = strides[1];
\end{lstlisting}
You may ask a View for its layout via its \lstinline!array_layout! typedef.
This can be helpful for C++ template metaprogramming.  For example:
\begin{lstlisting}
template<class ViewType>
void callBlas (const ViewType& A) {
  typedef typename ViewType::array_layout array_layout;
  if (std::is_same<array_layout, LayoutLeft>::value) {
    callSomeBlasFunction (A.ptr_on_device (), ...);
  } else {
    throw std::invalid_argument ("A is not LayoutLeft");
  }
}
\end{lstlisting}

@@

We discourage access to a View's ``raw'' pointer.  This circumvents
reference counting.  That is, the memory may be deallocated once the
View's reference count goes to zero, so holding on to a raw pointer
may result in invalid memory access.  Furthermore, it may not even be
possible to access the View's memory from a given execution space.
For example, a View in the \lstinline!Cuda! space points to CUDA device
memory. Also using raw pointers would normally defeat the usability 
of polymorpic layouts and automatic padding. 
Nevertheless, sometimes you really need access to the
pointer.  For such cases, we provide the \lstinline!ptr_on_device()!
method.  For example:
\begin{lstlisting}
// Legacy function that takes a raw pointer.
extern void legacyFunction (double* x_raw, const size_t len);

// Your function that takes a View.
void myFunction (const View<double*>& x) {
  // DON'T DO THIS UNLESS YOU MUST
  double* a_raw = a.ptr_on_device ();
  const size_t N = x.dimension_0 ();
  legacyFunction (a_raw, N);
}
\end{lstlisting}
A user is in most cases also allowed to obtain a pointer to a specific 
element via the usual \lstinline|&| operator. For example

@@

>>>>

Another important thing to keep in mind is that a \lstinline|View| handle is a 
stateful object. It is not legal to create a \lstinline|View| handle from
raw memory by typecasting a pointer. To call any operator on a \lstinline|View|
its constructor must have been called before. This includes the assignment
operator. If it is necessary to initialize raw memory with a \lstinline|View| 
handle, one can legally do so using a move constructor (``placement new'').
The above has nothing to do with the data a \lstinline|View| is referencing.
It is completely legal to give a typecast pointer to the constructor of an
unmanaged \lstinline|View|.

\begin{lstlisting}
View<int*> * a_ptr = (View<int*>*) malloc(10*sizeof(View<int*);
a_ptr[0] = View<int*>("A0",1000); // This is illegal
new(&a_ptr[1]) View<int*>("A1",10000); // This is legal 
\end{lstlisting}

@@

Other layouts are possible.  For example, Kokkos has a ``tiled''
layout, where a tile's entries are stored contiguously (in either row-
or column-major order) and tiles have compile-time dimensions.  One
may also use Kokkos to implement Morton ordering or variants thereof.
In order to write a custom layout one has to define a new layout class and
specialise the \lstinline|ViewMapping| class for that layout. 
The \lstinline|ViewMapping| class implements the offset operator as well
as stride calculation for regular layouts. 
A good way to start such a customization is by copying the implementation of 
\lstinline|LayoutLeft| and its associated \lstinline|ViewMapping| specialization,
renaming the layout and then change the offset operator.

@@

// Get 'LDA' for BLAS / LAPACK
int strides[2]; // any integer type works in stride()
A.stride (strides);
const int LDA = strides[1];
\end{lstlisting}
You may ask a View for its layout via its \lstinline!array_layout! typedef.
This can be helpful for C++ template metaprogramming.  For example:
\begin{lstlisting}
template<class ViewType>
void callBlas (const ViewType& A) {
  typedef typename ViewType::array_layout array_layout;
  if (std::is_same<array_layout, LayoutLeft>::value) {
    callSomeBlasFunction (A.data(), ...);
  } else {
    throw std::invalid_argument ("A is not LayoutLeft");
  }
}
\end{lstlisting}

@@

We discourage access to a View's ``raw'' pointer.  This circumvents
reference counting.  That is, the memory may be deallocated once the
View's reference count goes to zero, so holding on to a raw pointer
may result in invalid memory access.  Furthermore, it may not even be
possible to access the View's memory from a given execution space.
For example, a View in the \lstinline!Cuda! space points to CUDA device
memory. Also using raw pointers would normally defeat the usability 
of polymorpic layouts and automatic padding. 
Nevertheless, sometimes you really need access to the
pointer.  For such cases, we provide the \lstinline!data()!
method.  For example:
\begin{lstlisting}
// Legacy function that takes a raw pointer.
extern void legacyFunction (double* x_raw, const size_t len);

// Your function that takes a View.
void myFunction (const View<double*>& x) {
  // DON'T DO THIS UNLESS YOU MUST
  double* a_raw = a.data ();
  const size_t N = x.dimension_0 ();
  legacyFunction (a_raw, N);
}
\end{lstlisting}
A user is in most cases also allowed to obtain a pointer to a specific 
element via the usual \lstinline|&| operator. For example

@@

\subsection{Conversion Rules and Function Specialization}

Not all view types can be assigned to each other. In summary
the data type and dimension have to match, the layout must be
compatible and the memory space has to match. Examples illustrating
the rules are:

\begin{enumerate}
\item Memory Spaces must match
\subitem \lstinline{ Kokkos::View<int*> -> Kokkos::View<int*,HostSpace> // ok if default memory space is HostSpace}
\item Data Type and Rank has to Match
\subitem \lstinline{int*  -> int*          //ok}
\subitem \lstinline{int*  -> const int*    //ok}
\subitem \lstinline{const int*  -> int*    //not ok, const violation}
\subitem \lstinline{int** -> int*          //not ok, rank mismatch}
\subitem \lstinline{int*[3] -> int**       //ok}
\subitem \lstinline{int** -> int*[3]       //ok if runtime dimension check matches}
\subitem \lstinline{int*  -> long*         //not ok, type mismatch}
\item Layouts must be compatible
\subitem \lstinline{LayoutRight -> LayoutRight //ok}
\subitem \lstinline{LayoutLeft -> LayoutRight  //not ok}
\subitem \lstinline{LayoutLeft -> LayoutSride  //ok}
\subitem \lstinline{LayoutStride -> LayoutLeft //ok if runtime dimensions allow assignment}
\item Memory Traits 
\end{enumerate}
